### 2019-02-20 post-linearity-2018/final-submission/LAM-output.tex
Final version sources.

Similar as the sequent calculus, the term assignment for CNC logic is
also composed of two logics; intuitionistic linear logic on the left,
denoted by $\cat{C}$, and the Lambek calculus on the right, denoted by
$\cat{L}$. The syntax for types and contexts we use in the term
assignment is the same as in the sequent calculus. The rest of the
syntax for the term assignment is defined as follows.
\begin{definition}
  \label{def:Lambek-syntax}
  The following grammar describes the syntax of the term assignment of the
  CNC logic:
  \begin{center}\vspace{-3px}\small
    \begin{math}
      \begin{array}{lll}        
        \text{($\cat{C}$-Terms)} & [[t]] ::= [[x]] \mid [[trivT]] \mid [[t1 (*) t2]] \mid [[let t1 : X be q in t2]] \mid [[\x:X.t]] \mid [[t1 t2]] \mid [[ex t1 , t2 with x1 , x2 in t3]] \mid [[Gf s]]\\
        \text{($\cat{L}$-Terms)} & [[s]] ::= [[x]] \mid [[trivS]] \mid [[s1 (>) s2]] \mid [[let s1 : A be p in s2]] \mid [[let t : X be q in s]] \mid [[\l x : A.s]] \mid [[\r x : A . s]] \\
        & \,\,\,\,\,\,\,\,\,\mid [[appl s1 s2]] \mid [[appr s1 s2]] \mid [[F t]]\\        
        \text{($\cat{C}$-Patterns)} & [[q]] ::= [[trivTp]] \mid [[x]] \mid [[q1 (*) q2]] \mid [[Gf p]]\\
        \text{($\cat{L}$-Patterns)} & [[p]] ::= [[trivSp]] \mid [[x]] \mid [[p1 (>) p2]] \mid [[F q]]\\        
        \text{($\cat{C}$-Typing Judgment)} & [[I |-c t : X]]\\
        \text{($\cat{L}$-Typing Judgment)} & [[G |-l s : A]]\\
      \end{array}
    \end{math}
  \end{center}
\end{definition}

Now $\cat{C}$-typing judgments are denoted by $[[P |-c t : X]]$ where
$[[P]]$ is a sequence of pairs of variables and their types, denoted by
$[[x : X]]$, $[[t]]$ is a $\cat{C}$-term, and $[[X]]$ is a $\cat{C}$-type.  
The $\cat{C}$-terms are all standard, but $[[Gf s]]$ corresponds to the
morphism part of the right-adjoint of the adjunction between both logics,
and $[[ex t1 , t2 with x1 , x2 in t3]]$ is the introduction form for the
structural rule exchange.

The $\cat{L}$-typing judgment has the form $[[G |-l s : A]]$ where $[[G]]$
is now a $\cat{L}$-context, denoted by $[[G]]$ or $[[D]]$. These contexts
are ordered sequences of pairs of free variables with their types from
\emph{both} sides denoted by $[[x : B]]$ and $[[x : X]]$ respectively.
Finally, the term $[[s]]$ is a $\cat{L}$-term, and $[[A]]$ is a
$\cat{L}$-type.  Given two typing contexts $[[G]]$ and $[[D]]$ we denote
their concatenation by $[[G;D]]$; we use a semicolon here to emphasize the
fact that the contexts are ordered. $\cat{L}$-terms correspond to
introduction and elimination forms for each of the previous types. For
example, $[[s1 (>) s2]]$ introduces a tensor, and
$[[let s1 : A (>) B be x (>) y in s2]]$ eliminates a tensor.

The typing rules for CNC logic can be found in
Figure~\ref{fig:CNC-typing-rules}.
\begin{figure}
  \footnotesize
  \begin{tabular}{|c|}
    \hline\\
      \begin{mathpar}
      \NDdruleTXXid{} \and
      \NDdruleTXXunitI{} \and
      \NDdruleTXXunitE{} \and
      \NDdruleTXXtenI{} \and
      \NDdruleTXXtenE{} \and
      \NDdruleTXXimpI{} \and
      \NDdruleTXXimpE{} \and
      \NDdruleTXXGI{} \and
      \NDdruleTXXbeta{} \and
      \NDdruleTXXcut{}      
      \end{mathpar}
      \\
      \\
      \hline
      \\[5px]
    \begin{mathpar}
      \NDdruleSXXid{} \and
      \NDdruleSXXunitI{} \and
      \NDdruleSXXunitETwo{} \and
      \NDdruleSXXunitEOne{} \and
      \NDdruleSXXtenI{} \and
      \NDdruleSXXtenETwo{} \and
      \NDdruleSXXtenEOne{} \and
      \NDdruleSXXimprI{} \and
      \NDdruleSXXimprE{} \and
      \NDdruleSXXimplI{} \and
      \NDdruleSXXimplE{} \and
      \NDdruleSXXFI{} \and
      \NDdruleSXXFE{} \and
      \NDdruleSXXGE{} \and
      \NDdruleSXXbeta{} \and
      \NDdruleSXXcutTwo{} \and
      \NDdruleSXXcutOne{}
    \end{mathpar}\\\\
    \hline
  \end{tabular}  
  \caption{Typing Rules for CNC Logic}
  \label{fig:CNC-typing-rules}
\end{figure}
We split the figure in two: the top of the figure are the rules of
intuitionistic linear logic whose judgment is the $\mathcal{C}$-typing
judgment denoted by $[[P |-c t : X]]$, and the bottom of the figure
are the rules for the mixed commutative/non-commutative Lambek
calculus whose judgment is the $\mathcal{L}$-judgment denoted by
$[[G |-l s : A]]$, and the two halves are connected via the rules 
rules $\NDdruleTXXGIName{}$, $\NDdruleSXXGEName{}$,
$\NDdruleSXXFIName{}$, and $\NDdruleSXXFEName{}$,
$\NDdruleSXXunitEOneName{}$, $\NDdruleSXXtenEOneName{}$, and
$\NDdruleSXXcutOneName{}$.

@@

We also proved that the sequent calculus formalization given in
Figure~\ref{fig:CNC-sequent-calculus} is equivalent to the typing rules (or
else called the natural deduction formalization) given in
Figure~\ref{fig:CNC-typing-rules} are equivalent, as stated in the following
theorem.
\begin{theorem}
  \label{thm:sc-nd-equiv}
  The sequent calculus ($\mathit{SC}$) and natural deduction ($\mathit{ND}$)
  formalizations for CNC logic are equivalent in the sense that there are
  two mappings $N:\mathit{SC}\rightarrow\mathit{ND}$ and
  $S:\mathit{ND}\rightarrow\mathit{SC}$ that map each rule in $\mathit{SC}$
  to a proof in $\mathit{ND}$, and each rule in $\mathit{ND}$ to a proof
  in $\mathit{SC}$, respectively.
\end{theorem}
\begin{proof}
  The proof is done case by case on each rule in the sequence calculus and
  natural deduction formalizations. It is obvious that the axioms in one
  formalization can be mapped to the axioms in the other. The introduction
  rules in $\mathit{ND}$ are mapped to the right rules in $\mathit{SC}$, and
  vice versa. The elimination rules and lefts rules are mapped to each other
  with some fiddling. For instance, the elimination rule for the
  non-commutative tensor is mapped to the following proof in $\mathit{SC}$:
  \begin{center}
    \scriptsize
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\scriptsize $\ElledruleTXXcutName$}] {
        {[[I |-c t1 : X (*) Y]]} \\
        $$\mprset{flushleft}
        \inferrule* [right={\scriptsize $\ElledruleTXXtenLName$}] {
          {[[P1, x : X, y : Y, P2 |-c t2 : Z]]}
        }{[[P1, z : X (*) Y, P2 |-c let z : X (*) Y be x (*) y in t2 : Z]]}
      }{[[P1, I, P2 |-c [t1 / z](let z : X (*) Y be x (*) y in t2) : Z]]}
    \end{math}
  \end{center}
  The full proof is in Appendix~\ref{app:sc-nd-equiv}.
\end{proof}

%% \noindent
%% Finally, we have strong normalization of CNC logic.
%% 
%% \begin{theorem}[Strong Normalization]
%%   \label{thm:strong-normalization}
%%   CNC logic is strongly normalizing.
%% \end{theorem}
%% \begin{proof}
%%   The proof is via a translation to LNL logic, which has been proved to be
%%   strongly normalizing \cite{Mellies:2009}. The proof is in
%%   Appendix~\ref{app:strong-normalization}. We first define a function
%%   $CL$ that maps our term assignment for natural deduction to Benton's
%%   term assignment, and then shows that if there is reduction $[[t1 ~>  t2]]$ (resp. $[[s1 ~> s2]]$) in CNC, then there is a reduction
%%   $CL(t_1)\leadsto CL(t_2)$ (resp. $CL(s_1)\leadsto CL(s_2)$) in
%%   LNL. Recall that a LNL model consists of an adjunction
%%   $F:\cat{C}\dashv\cat{L}:G$ in which $\cat{C}$ is a Cartesian closed
%%   category and $\cat{L}$ is a SMCC.  The basic idea of mapping a LAM
%%   model to a LNL model is mapping the SMCC in LAM to the Cartesian
%%   closed category in LNL, and the Lambek category in LAM to the SMCC in
%%   LNL. For instance, $s_1\triangleright s_2$ in CNC, where $[[s1]]$ and
%%   $[[s2]]$ are atomic terms, maps to $CL(s_1)\otimes CL(s_2)$ in LNL.
%% \end{proof}

>>>>

Similar as the sequent calculus, the term assignment for CNC logic is
also composed of two logics; intuitionistic linear logic on the left,
denoted by $\cat{C}$, and the Lambek calculus on the right, denoted by
$\cat{L}$. The syntax for types and contexts we use in the term
assignment is the same as in the sequent calculus. The rest of the
syntax for the term assignment is defined as follows.
\begin{definition}
  \label{def:Lambek-syntax}
  The following grammar describes the syntax of the term assignment of the
  CNC logic:
  \begin{center}\vspace{-3px}\small
    \begin{math}
      \begin{array}{lll}        
        \text{($\cat{C}$-Terms)} & \NDnt{t} ::= \NDmv{x} \mid  \mathsf{triv}  \mid \NDnt{t_{{\mathrm{1}}}}  \otimes  \NDnt{t_{{\mathrm{2}}}} \mid  \mathsf{let}\, \NDnt{t_{{\mathrm{1}}}}  :  \NDnt{X} \,\mathsf{be}\, \NDnt{q} \,\mathsf{in}\, \NDnt{t_{{\mathrm{2}}}}  \mid  \lambda  \NDmv{x}  :  \NDnt{X} . \NDnt{t}  \mid  \NDnt{t_{{\mathrm{1}}}}   \NDnt{t_{{\mathrm{2}}}}  \mid  \mathsf{ex}\, \NDnt{t_{{\mathrm{1}}}} , \NDnt{t_{{\mathrm{2}}}} \,\mathsf{with}\, \NDmv{x_{{\mathrm{1}}}} , \NDmv{x_{{\mathrm{2}}}} \,\mathsf{in}\, \NDnt{t_{{\mathrm{3}}}}  \mid  \mathsf{G}\, \NDnt{s} \\
        \text{($\cat{L}$-Terms)} & \NDnt{s} ::= \NDmv{x} \mid  \mathsf{triv}  \mid \NDnt{s_{{\mathrm{1}}}}  \triangleright  \NDnt{s_{{\mathrm{2}}}} \mid  \mathsf{let}\, \NDnt{s_{{\mathrm{1}}}}  :  \NDnt{A} \,\mathsf{be}\, \NDnt{p} \,\mathsf{in}\, \NDnt{s_{{\mathrm{2}}}}  \mid  \mathsf{let}\, \NDnt{t}  :  \NDnt{X} \,\mathsf{be}\, \NDnt{q} \,\mathsf{in}\, \NDnt{s}  \mid  \lambda_l  \NDmv{x}  :  \NDnt{A} . \NDnt{s}  \mid  \lambda_r  \NDmv{x}  :  \NDnt{A} . \NDnt{s}  \\
        & \,\,\,\,\,\,\,\,\,\mid  \mathsf{app}_l\, \NDnt{s_{{\mathrm{1}}}} \, \NDnt{s_{{\mathrm{2}}}}  \mid  \mathsf{app}_r\, \NDnt{s_{{\mathrm{1}}}} \, \NDnt{s_{{\mathrm{2}}}}  \mid  \mathsf{F} \NDnt{t} \\        
        \text{($\cat{C}$-Patterns)} & \NDnt{q} ::=  \mathsf{triv}  \mid \NDmv{x} \mid \NDnt{q_{{\mathrm{1}}}}  \otimes  \NDnt{q_{{\mathrm{2}}}} \mid  \mathsf{G}\, \NDnt{p} \\
        \text{($\cat{L}$-Patterns)} & \NDnt{p} ::=  \mathsf{triv}  \mid \NDmv{x} \mid \NDnt{p_{{\mathrm{1}}}}  \triangleright  \NDnt{p_{{\mathrm{2}}}} \mid  \mathsf{F}\, \NDnt{q} \\        
        \text{($\cat{C}$-Typing Judgment)} & \Phi  \vdash_\mathcal{C}  \NDnt{t}  \NDsym{:}  \NDnt{X}\\
        \text{($\cat{L}$-Typing Judgment)} & \Gamma  \vdash_\mathcal{L}  \NDnt{s}  \NDsym{:}  \NDnt{A}\\
      \end{array}
    \end{math}
  \end{center}
\end{definition}

Now $\cat{C}$-typing judgments are denoted by $\Psi  \vdash_\mathcal{C}  \NDnt{t}  \NDsym{:}  \NDnt{X}$ where
$\Psi$ is a sequence of pairs of variables and their types, denoted by
$\NDmv{x}  \NDsym{:}  \NDnt{X}$, $\NDnt{t}$ is a $\cat{C}$-term, and $\NDnt{X}$ is a $\cat{C}$-type.  
The $\cat{C}$-terms are all standard, but $ \mathsf{G}\, \NDnt{s} $ corresponds to the
morphism part of the right-adjoint of the adjunction between both logics,
and $ \mathsf{ex}\, \NDnt{t_{{\mathrm{1}}}} , \NDnt{t_{{\mathrm{2}}}} \,\mathsf{with}\, \NDmv{x_{{\mathrm{1}}}} , \NDmv{x_{{\mathrm{2}}}} \,\mathsf{in}\, \NDnt{t_{{\mathrm{3}}}} $ is the introduction form for the
structural rule exchange.

The $\cat{L}$-typing judgment has the form $\Gamma  \vdash_\mathcal{L}  \NDnt{s}  \NDsym{:}  \NDnt{A}$ where $\Gamma$
is now a $\cat{L}$-context, denoted by $\Gamma$ or $\Delta$. These contexts
are ordered sequences of pairs of free variables with their types from
\emph{both} sides denoted by $\NDmv{x}  \NDsym{:}  \NDnt{B}$ and $\NDmv{x}  \NDsym{:}  \NDnt{X}$ respectively.
Finally, the term $\NDnt{s}$ is a $\cat{L}$-term, and $\NDnt{A}$ is a
$\cat{L}$-type.  Given two typing contexts $\Gamma$ and $\Delta$ we denote
their concatenation by $\Gamma  \NDsym{;}  \Delta$; we use a semicolon here to emphasize the
fact that the contexts are ordered. $\cat{L}$-terms correspond to
introduction and elimination forms for each of the previous types. For
example, $\NDnt{s_{{\mathrm{1}}}}  \triangleright  \NDnt{s_{{\mathrm{2}}}}$ introduces a tensor, and
$ \mathsf{let}\, \NDnt{s_{{\mathrm{1}}}}  :  \NDnt{A}  \triangleright  \NDnt{B} \,\mathsf{be}\, \NDmv{x}  \triangleright  \NDmv{y} \,\mathsf{in}\, \NDnt{s_{{\mathrm{2}}}} $ eliminates a tensor.

The typing rules for CNC logic can be found in
Figure~\ref{fig:CNC-typing-rules}.
\begin{figure}
  \footnotesize
  \begin{tabular}{|c|}
    \hline\\
      \begin{mathpar}
      \NDdruleTXXid{} \and
      \NDdruleTXXunitI{} \and
      \NDdruleTXXunitE{} \and
      \NDdruleTXXtenI{} \and
      \NDdruleTXXtenE{} \and
      \NDdruleTXXimpI{} \and
      \NDdruleTXXimpE{} \and
      \NDdruleTXXGI{} \and
      \NDdruleTXXbeta{} \and
      \NDdruleTXXcut{}      
      \end{mathpar}
      \\
      \\
      \hline
      \\[5px]
    \begin{mathpar}
      \NDdruleSXXid{} \and
      \NDdruleSXXunitI{} \and
      \NDdruleSXXunitETwo{} \and
      \NDdruleSXXunitEOne{} \and
      \NDdruleSXXtenI{} \and
      \NDdruleSXXtenETwo{} \and
      \NDdruleSXXtenEOne{} \and
      \NDdruleSXXimprI{} \and
      \NDdruleSXXimprE{} \and
      \NDdruleSXXimplI{} \and
      \NDdruleSXXimplE{} \and
      \NDdruleSXXFI{} \and
      \NDdruleSXXFE{} \and
      \NDdruleSXXGE{} \and
      \NDdruleSXXbeta{} \and
      \NDdruleSXXcutTwo{} \and
      \NDdruleSXXcutOne{}
    \end{mathpar}\\\\
    \hline
  \end{tabular}  
  \caption{Typing Rules for CNC Logic}
  \label{fig:CNC-typing-rules}
\end{figure}
We split the figure in two: the top of the figure are the rules of
intuitionistic linear logic whose judgment is the $\mathcal{C}$-typing
judgment denoted by $\Psi  \vdash_\mathcal{C}  \NDnt{t}  \NDsym{:}  \NDnt{X}$, and the bottom of the figure
are the rules for the mixed commutative/non-commutative Lambek
calculus whose judgment is the $\mathcal{L}$-judgment denoted by
$\Gamma  \vdash_\mathcal{L}  \NDnt{s}  \NDsym{:}  \NDnt{A}$, and the two halves are connected via the rules 
rules $\NDdruleTXXGIName{}$, $\NDdruleSXXGEName{}$,
$\NDdruleSXXFIName{}$, and $\NDdruleSXXFEName{}$,
$\NDdruleSXXunitEOneName{}$, $\NDdruleSXXtenEOneName{}$, and
$\NDdruleSXXcutOneName{}$.

@@

We also proved that the sequent calculus formalization given in
Figure~\ref{fig:CNC-sequent-calculus} is equivalent to the typing rules (or
else called the natural deduction formalization) given in
Figure~\ref{fig:CNC-typing-rules} are equivalent, as stated in the following
theorem.
\begin{theorem}
  \label{thm:sc-nd-equiv}
  The sequent calculus ($\mathit{SC}$) and natural deduction ($\mathit{ND}$)
  formalizations for CNC logic are equivalent in the sense that there are
  two mappings $N:\mathit{SC}\rightarrow\mathit{ND}$ and
  $S:\mathit{ND}\rightarrow\mathit{SC}$ that map each rule in $\mathit{SC}$
  to a proof in $\mathit{ND}$, and each rule in $\mathit{ND}$ to a proof
  in $\mathit{SC}$, respectively.
\end{theorem}
\begin{proof}
  The proof is done case by case on each rule in the sequence calculus and
  natural deduction formalizations. It is obvious that the axioms in one
  formalization can be mapped to the axioms in the other. The introduction
  rules in $\mathit{ND}$ are mapped to the right rules in $\mathit{SC}$, and
  vice versa. The elimination rules and lefts rules are mapped to each other
  with some fiddling. For instance, the elimination rule for the
  non-commutative tensor is mapped to the following proof in $\mathit{SC}$:
  \begin{center}
    \scriptsize
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\scriptsize $\ElledruleTXXcutName$}] {
        {\Phi  \vdash_\mathcal{C}  \NDnt{t_{{\mathrm{1}}}}  \NDsym{:}  \NDnt{X}  \otimes  \NDnt{Y}} \\
        $$\mprset{flushleft}
        \inferrule* [right={\scriptsize $\ElledruleTXXtenLName$}] {
          {\Psi_{{\mathrm{1}}}  \NDsym{,}  \NDmv{x}  \NDsym{:}  \NDnt{X}  \NDsym{,}  \NDmv{y}  \NDsym{:}  \NDnt{Y}  \NDsym{,}  \Psi_{{\mathrm{2}}}  \vdash_\mathcal{C}  \NDnt{t_{{\mathrm{2}}}}  \NDsym{:}  \NDnt{Z}}
        }{\Psi_{{\mathrm{1}}}  \NDsym{,}  \NDmv{z}  \NDsym{:}  \NDnt{X}  \otimes  \NDnt{Y}  \NDsym{,}  \Psi_{{\mathrm{2}}}  \vdash_\mathcal{C}   \mathsf{let}\, \NDmv{z}  :  \NDnt{X}  \otimes  \NDnt{Y} \,\mathsf{be}\, \NDmv{x}  \otimes  \NDmv{y} \,\mathsf{in}\, \NDnt{t_{{\mathrm{2}}}}   \NDsym{:}  \NDnt{Z}}
      }{\Psi_{{\mathrm{1}}}  \NDsym{,}  \Phi  \NDsym{,}  \Psi_{{\mathrm{2}}}  \vdash_\mathcal{C}  \NDsym{[}  \NDnt{t_{{\mathrm{1}}}}  \NDsym{/}  \NDmv{z}  \NDsym{]}  \NDsym{(}   \mathsf{let}\, \NDmv{z}  :  \NDnt{X}  \otimes  \NDnt{Y} \,\mathsf{be}\, \NDmv{x}  \otimes  \NDmv{y} \,\mathsf{in}\, \NDnt{t_{{\mathrm{2}}}}   \NDsym{)}  \NDsym{:}  \NDnt{Z}}
    \end{math}
  \end{center}
  The full proof is in Appendix~\ref{app:sc-nd-equiv}.
\end{proof}

%% \noindent
%% Finally, we have strong normalization of CNC logic.
%% 
%% \begin{theorem}[Strong Normalization]
%%   \label{thm:strong-normalization}
%%   CNC logic is strongly normalizing.
%% \end{theorem}
%% \begin{proof}
%%   The proof is via a translation to LNL logic, which has been proved to be
%%   strongly normalizing \cite{Mellies:2009}. The proof is in
%%   Appendix~\ref{app:strong-normalization}. We first define a function
%%   $CL$ that maps our term assignment for natural deduction to Benton's
%%   term assignment, and then shows that if there is reduction $\NDnt{t_{{\mathrm{1}}}}  \leadsto_\beta  \NDnt{t_{{\mathrm{2}}}}$ (resp. $\NDnt{s_{{\mathrm{1}}}}  \leadsto_\beta  \NDnt{s_{{\mathrm{2}}}}$) in CNC, then there is a reduction
%%   $CL(t_1)\leadsto CL(t_2)$ (resp. $CL(s_1)\leadsto CL(s_2)$) in
%%   LNL. Recall that a LNL model consists of an adjunction
%%   $F:\cat{C}\dashv\cat{L}:G$ in which $\cat{C}$ is a Cartesian closed
%%   category and $\cat{L}$ is a SMCC.  The basic idea of mapping a LAM
%%   model to a LNL model is mapping the SMCC in LAM to the Cartesian
%%   closed category in LNL, and the Lambek category in LAM to the SMCC in
%%   LNL. For instance, $s_1\triangleright s_2$ in CNC, where $\NDnt{s_{{\mathrm{1}}}}$ and
%%   $\NDnt{s_{{\mathrm{2}}}}$ are atomic terms, maps to $CL(s_1)\otimes CL(s_2)$ in LNL.
%% \end{proof}

### 2019-02-20 post-linearity-2018/final-submission/adjoint-models.tex
Final version sources.

In this section we introduce Lambek Adjoint Models (LAMs). Benton's
LNL model consists of a symmetric monoidal adjunction
$F:\cat{C}\dashv\cat{L}:G$ between a cartesian closed category
$\cat{C}$ and a symmetric monoidal closed category $\cat{L}$. LAM
consists of a monoidal adjunction between a symmetric monoidal closed
category and a Lambek category.
\begin{definition}
  \label{def:lambek-category}
  A \textbf{Lambek category} is a monoidal category $(\cat{L},\tri,I',\alpha',\lambda',\rho')$
  with two functors $- \rightharpoonup - : \cat{L}^{\mathsf{op}} \times \cat{L} \mto \cat{L}$ and
  $- \leftharpoonup - : \cat{L} \times \cat{L}^{\mathsf{op}} \mto \cat{L}$ such that the following
  two natural bijections hold:
  \[
  \begin{array}{lllll}
    \Hom{L}{A \tri B}{C} \cong \Hom{L}{A}{B \rightharpoonup C} & \quad &
    \Hom{L}{A \tri B}{C} \cong \Hom{L}{B}{C \leftharpoonup A}\\
  \end{array}
  \]  
\end{definition}
\noindent
Lambek categories are also known as monoidal bi-closed categories.

@@

\textbf{Strong non-commutative monad}. Next we show that the monad on
$\cat{C}$ in LAM is strong but non-commutative. In Benton's LNL model,
the monad on the cartesian closed category is commutative, but later
Benton and Wadler \cite{Benton:1996} wonder, is it possible to model
non-commutative monads using adjoint models similar to LNL models? The
following answers their question in the positive w.r.t. the monoidal monads.
\begin{lemma}
\label{lem:monoidal-monad}
The monad induced by any LAM, $GF : \cat{C} \mto \cat{C}$, is monoidal.
\end{lemma}
\begin{proof}
  The proof is done by checking the conditions for a functor being monoidal.
  The detail of the proof is in Appendix~\ref{app:monoidal-monad}.
\end{proof}
\noindent
However, the monad is not symmetric because the following diagram does
not commute.
\begin{mathpar}
\bfig
  \ptriangle/->`->`/<900,400>[
    GFX\otimes GFY`GFY\otimes GFX`G(FX\tri FY);\e{GFX,GFY}`\n{FX,FY}`]
  \morphism(900,400)<900,0>[GFY\otimes GFX`G(FY\tri FX);\n{FY,FX}]
  \dtriangle(900,0)/`->`->/<900,400>[
    G(FY\tri FX)`GF(X\otimes Y)`GF(Y\otimes X);`G\m{Y,X}`GF\e{X,Y}]
  \morphism|b|<900,0>[G(FX\tri FY)`GF(X\otimes Y);G\m{X,Y}]
\efig
\end{mathpar}
Commutativity fails, because the functors defining the monad are not
symmetric monoidal, but only monoidal. This means that the diagram
\[
\bfig
\square<700,400>[
  FA\otimes'FB`FB\otimes'FA`F(A\otimes B)`F(B\otimes A);
  \e{FA,FB}`\m{A,B}`\m{B,A}`F\e{A,B}]
\efig
\]
does not hold for $G$ nor $F$.  However, we can prove the monad is
strong.
\begin{lemma}
  \label{lem:strong-monad}
  The monad, $GF : \cat{C} \mto \cat{C}$, on the symmetric monoidal
  closed category in LAM is strong.
\end{lemma}
\begin{proof}
The proof is done by frist defining a natural transformation $\tau$, called
the \textbf{tensorial strength}, with components
$\tau_{A,B}:A\tri TB\rightarrow T(A\tri B)$, and then proving the
commutativity of several diagrams through diagram chasing. The formal
definition for a strong monad and the full proof are in
Appendix~\ref{app:strong-monad}.
\end{proof}
\noindent
Finally, we obtain the non-communativity of the monad induced by some 
LAM as follows.
\begin{lemma}[Due to Kock~\cite{kock1972strong}]
\label{lem:monad-com-iff-sym}
  Let $\cat{M}$ be a symmetric monoidal category and $T$ be a strong monad on $\cat{M}$. Then
  $T$ is commutative iff it is symmetric monoidal.
\end{lemma}

@@

\textbf{Comonad for exchange}.  We conclude this section by showing
that the comonad induced by some LAM is monoidal and extends $\cat{L}$
with exchange. The former is proved in \cite{kelly1974doctrinal}. The
latter is shown by proving that its corresponding co-Eilenberg-Moore
category is symmetric monoidal closed.

@@

\begin{corollary}
  The subcategory, $\mathsf{Exp}(\cat{L}^{FG})$, of the
  co-Eilenberg-Moore category $\cat{L}^{FG}$ consisting of the free
  exponential coalgebras is symmetric monoidal closed.
\end{corollary}
\begin{proof}
  The proof that $\mathsf{Exp}(\cat{L}^{FG})$ is monoidal closed
  follows similarly to Benton~\cite{Benton:1994}.
\end{proof}

>>>>

In this section we introduce Lambek Adjoint Models (LAMs). Benton's
LNL model consists of a symmetric monoidal adjunction
$F:\cat{C}\dashv\cat{L}:G$ between a Cartesian closed category
$\cat{C}$ and a symmetric monoidal closed category $\cat{L}$. LAM
consists of a monoidal adjunction between a symmetric monoidal closed
category and a Lambek category.
\begin{definition}
  \label{def:lambek-category}
  A \textbf{Lambek category} is a monoidal category $(\cat{L},\tri,I',\alpha',\lambda',\rho')$
  with two functors $- \rightharpoonup - : \cat{L}^{\mathsf{op}} \times \cat{L} \mto \cat{L}$ and
  $- \leftharpoonup - : \cat{L} \times \cat{L}^{\mathsf{op}} \mto \cat{L}$ such that the following
  two natural bijections hold:
  \[
  \begin{array}{lllll}
    \Hom{L}{A \tri B}{C} \cong \Hom{L}{A}{B \rightharpoonup C} & \quad &
    \Hom{L}{A \tri B}{C} \cong \Hom{L}{B}{C \leftharpoonup A}\\
  \end{array}
  \]  
\end{definition}
\noindent
Lambek categories are also known as monoidal bi-closed categories.

@@

\textbf{Strong non-commutative monad}. Next we show that the monad on
$\cat{C}$ in LAM is strong but non-commutative. In Benton's LNL model,
the monad on the Cartesian closed category is commutative, but later
Benton and Wadler \cite{Benton:1996} wonder, is it possible to model
non-commutative monads using adjoint models similar to LNL models? The
following shows that LAMs correspond to strong non-commutative
monoidal monads.
\begin{lemma}
\label{lem:monoidal-monad}
The monad induced by any LAM, $GF : \cat{C} \mto \cat{C}$, is monoidal.
\end{lemma}
\begin{proof}
  The proof is done by checking the conditions for a functor being monoidal.
  The detail of the proof is in Appendix~\ref{app:monoidal-monad}.
\end{proof}
\noindent
However, the monad is not symmetric because the following diagram does
not commute.
\begin{mathpar}
\bfig
  \ptriangle/->`->`/<900,400>[
    GFX\otimes GFY`GFY\otimes GFX`G(FX\tri FY);\e{GFX,GFY}`\n{FX,FY}`]
  \morphism(900,400)<900,0>[GFY\otimes GFX`G(FY\tri FX);\n{FY,FX}]
  \dtriangle(900,0)/`->`->/<900,400>[
    G(FY\tri FX)`GF(X\otimes Y)`GF(Y\otimes X);`G\m{Y,X}`GF\e{X,Y}]
  \morphism|b|<900,0>[G(FX\tri FY)`GF(X\otimes Y);G\m{X,Y}]
\efig
\end{mathpar}
Commutativity fails, because the functors defining the monad are not
symmetric monoidal, but only monoidal. This means that the diagram
\[
\bfig
\square<700,400>[
  FA\otimes'FB`FB\otimes'FA`F(A\otimes B)`F(B\otimes A);
  \e{FA,FB}`\m{A,B}`\m{B,A}`F\e{A,B}]
\efig
\]
does not hold for $G$ nor $F$.  However, we can prove the monad is
strong.
\begin{lemma}
  \label{lem:strong-monad}
  The monad, $GF : \cat{C} \mto \cat{C}$, on the symmetric monoidal
  closed category in LAM is strong.
\end{lemma}
\begin{proof}
The proof is done by first defining a natural transformation $\tau$, called
the \textbf{tensorial strength}, with components
$\tau_{A,B}:A\tri TB\rightarrow T(A\tri B)$, and then proving the
commutativity of several diagrams through diagram chasing. The formal
definition for a strong monad and the full proof are in
Appendix~\ref{app:strong-monad}.
\end{proof}
\noindent
Finally, we obtain the non-commutativity of the monad induced by some 
LAM as follows.
\begin{lemma}[Due to Kock~\cite{kock1972strong}]
\label{lem:monad-com-iff-sym}
  Let $\cat{M}$ be a symmetric monoidal category and $T$ be a strong monad on $\cat{M}$. Then
  $T$ is commutative iff it is symmetric monoidal.
\end{lemma}

@@

\textbf{Comonad for exchange}.  We conclude this section by showing
that the comonad induced by some LAM is monoidal and extends $\cat{L}$
with exchange. The former is proved in \cite{kelly1974doctrinal}. The
latter is shown by proving that its corresponding co-Eilenberg-Moore
category is symmetric monoidal.

@@

%% \begin{corollary}
%%   The subcategory, $\mathsf{Exp}(\cat{L}^{FG})$, of the
%%   co-Eilenberg-Moore category $\cat{L}^{FG}$ consisting of the free
%%   exponential coalgebras is symmetric monoidal closed.
%% \end{corollary}
%% \begin{proof}
%%   The proof that $\mathsf{Exp}(\cat{L}^{FG})$ is monoidal closed
%%   follows similarly to Benton~\cite{Benton:1994}.
%% \end{proof}

### 2019-02-20 post-linearity-2018/final-submission/fwork.tex
Final version sources.

Hasegawa~\cite{EPTCS238.6} studies the linear of-course modality,
$!A$, as a comonad induced by an adjunction between a cartesian closed
category a (non-symmetric) monoidal category.  The results here
generalizes his by generalizing the cartesian closed category to a
symmetric monoidal closed category.  However, his approach focuses on
the comonad rather than the adjunctions.  It would be interesting to
do the same for LAM as well.

>>>>

Hasegawa~\cite{EPTCS238.6} studies the linear of-course modality,
$!A$, as a comonad induced by an adjunction between a Cartesian closed
category a (non-symmetric) monoidal category.  The results here
generalizes his by generalizing the Cartesian closed category to a
symmetric monoidal closed category.  However, his approach focuses on
the comonad rather than the adjunctions.  It would be interesting to
do the same for LAM as well.

### 2019-02-20 post-linearity-2018/final-submission/introduction.tex
Final version sources.

Joachim Lambek first introduced the Syntactic Calculus, now known as
the Lambek Calculus, in 1958 \cite{Lambek1958}.  Since then the Lambek
Calculus has largely been motivated by providing an explanation of the
mathematics of sentence structure, and can be found at the core of
Categorial Grammar; a term first used in the title of Bar-Hillel,
Gaifman and Shamir (1960), but categorical grammar began with
Ajdukiewicz (1935) quite a few years earlier. At the end of the
eighties the Lambek calculus and other systems of categorial grammars
were taken up by computational linguists as exemplified by
\cite{oehrle2012categorial,moortgat1988categorial,Barry:1991:PFS:977180.977215,hepple1990grammar}.

>>>>

Joachim Lambek first introduced the Syntactic Calculus, now known as
the Lambek Calculus, in 1958 \cite{Lambek1958}.  Since then the Lambek
Calculus has largely been motivated by providing an explanation of the
mathematics of sentence structure, and can be found at the core of
Categorical Grammar; a term first used in the title of Bar-Hillel,
Gaifman and Shamir (1960), but categorical grammar began with
Ajdukiewicz (1935) quite a few years earlier. At the end of the
eighties the Lambek calculus and other systems of categorical grammars
were taken up by computational linguists as exemplified by
\cite{oehrle2012categorial,moortgat1988categorial,Barry:1991:PFS:977180.977215,hepple1990grammar}.
